---
title: "Remember-R"
author: "Camilo Gálvez A."
output: html_document
---

# Libraries
```{r Libraries, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
library(tidyverse)
library(vegan)
library(clustsig)
library(Biostrings)
library(DECIPHER)
library(phyloseq)
library(ape)
library(dendextend)
library(heatmaply)
library(viridis)
library(htmlwidgets)
library(taxa)
library(metacoder)
library(ggvegan)
library(doParallel)
library(hrbrthemes)
```


# Data
```{r Data, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
data <- read.csv("https://raw.githubusercontent.com/cmlglvz/Remember-R/main/Data/abundancia_taxonomia.csv", header = TRUE, sep = ";", skip = 0, row.names = 1)
environ <- read.csv("https://raw.githubusercontent.com/cmlglvz/Remember-R/main/Data/environment.csv", header = TRUE, sep = ";", dec = ".", skip = 0, row.names = 1)
```


# Abundance: Raw, relative and transformed
```{r Abundance, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
short <- dplyr::filter(data,Total >= 20)
abund <- short[,c(3:36)] %>% t() %>% as.data.frame()
colnames(abund) <- short$ASV

#Benjamin Glasner
rltv.Otu.Table <- function(x) {
  x.Data.rltv <- NULL
  for (i in 1:dim(x)[1]) {
    x.Data.rltv <- rbind(x.Data.rltv, x[i,]/apply(x, 1, function(x) sum(x))[i])
  }
  rownames(x.Data.rltv) <- rownames(x)
  invisible(x.Data.rltv)
}

rltv <- rltv.Otu.Table(abund) %>% as.data.frame()
rowSums(rltv)

vrltv <- vegan::decostand(abund, method = "total")
rowSums(vrltv)

hell <- vegan::decostand(abund, method = "hellinger")
rowSums(hell)

#vegan::decostand(vrltv, method = "hellinger") %>% rowSums()
```


# SIMPROF
The similarity profile routine ([SIMPROF](https://sites.google.com/site/mb3gustame/hypothesis-tests/simprof); Clarke et al. 2008) is a hypothesis test which may be used during exploratory analyses. SIMPROF tests data, with no required a priori groupings (in contrast to e.g. ANOSIM), against the null hypothesis of "absence of structure". 
```{r SIMPROF, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
bray <- vegan::vegdist(hell, method = "bray", diag = TRUE) %>% as.matrix() %>% as.data.frame()
smprf <- clustsig::simprof(data = bray, method.cluster = "average", silent = FALSE, increment = 1000, undef.zero = TRUE)
simprof.plot(smprf)

#ver si esto es util
hellprof <- add_column(hell, 
                       .before = hell[,1], 
                       Site = c(rep("Site.1", 12),
                                rep("Site.3", 12), 
                                rep("Site.5", 10)
                                ), 
                       .name_repair = "minimal")
hellprof$Site <- factor(hellprof$Site, levels = unique(hellprof$Site))
```


# Sequences analysis
```{r FASTA, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
#nms <- paste(data$ASV, data$Species, sep = "_")
fasta <- data.frame(names = short$ASV, Sequences = all_of(short$Seq))
seqRFLP::dataframe2fas(fasta, "X:/Documents/Proyectos_R/Remember-R/Products/asv.fasta")
#Less than 1000 seqs = muscle.exe -align asv.fasta -output alignment.fasta
#More than 1000 seqs = muscle.exe -super5 asv.fasta -output alignment.afa
#it took 20 minutes in this case
```


# Alignment dendrogram
```{r alignment, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
multialign <- "/Documents/Proyectos_R/Remember-R/Products/alignment.fasta"
dbConn <- dbConnect(SQLite(), ":memory:")
Seqs2DB(multialign, type = "FASTA", dbFile = dbConn, "")
x <- dbGetQuery(dbConn, "select description from Seqs")$description
Add2DB(myData = data.frame(identifier = x, stringsAsFactors = FALSE), dbConn)
consensus <- IdConsensus(dbConn, threshold = 0.3, minInformation = 0.1)
distance.matrix <- DistanceMatrix(consensus, correction = "Jukes-Cantor", processors = NULL, verbose = TRUE)
dendrogram <- IdClusters(distance.matrix, method = "ML", showPlot = TRUE, type = "dendrogram", myXStringSet = consensus, processors = NULL, verbose = TRUE)
#Selected model was:TN93 + G4, it took 144.29 secs
dbDisconnect(dbConn)
```


# Cluster manipulation
```{r cluster, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
clust <- as.dendrogram(as.hclust(dendrogram)) %>% dendextend::set("branches_lwd", 0.3) %>% dendextend::ladderize(right = TRUE)
plot(clust)
sclust <- smprf$hclust
sclustS <- as.dendrogram(sclust) %>% set("branches_lwd", 0.3) %>% ladderize(right = FALSE)
plot(sclustS)
```


# Heatmap
```{r heatmap, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
maligned <- readDNAMultipleAlignment(multialign, format = "fasta")
mstring <- as(maligned, "DNAStringSet")
DECIPHER::BrowseSeqs(mstring, htmlFile = "/Documents/Proyectos_R/Remember-R/Products/multiple_alignment.html")

#Check if ID from .fasta and abundance table match
print(all(colnames(hell)%in%gtools::mixedsort(as.vector(rownames(as.matrix(maligned))), decreasing = FALSE)))
#If TRUE = they match

heatmap <- heatmaply(normalize(hell), 
                     Colv = clust, 
                     Rowv = sclustS, 
                     main = "Here goes the title", 
                     color = viridis(n = 256, 
                                     alpha = 1, 
                                     begin = 0, 
                                     end = 1, 
                                     option = "rocket")
                     )
htmlwidgets::saveWidget(heatmap, file = "/Documents/Proyectos_R/Remember-R/Products/heatmap.html")

wos.heat <- heatmaply(normalize(hell), 
                      Colv = clust, 
                      Rowv = NA, 
                      main = "Heatmap but without SIMPROF", 
                      color = viridis(n = 256, 
                                      alpha = 0.7, 
                                      begin = 0, 
                                      end = 1, 
                                      option = "mako")
                      )

saveWidget(wos.heat, file = "/Documents/Proyectos_R/Remember-R/Products/wos_heat.html")
```


# Using phyloseq
This next "chapter" is based on Daniel Vaulot's online tutorial for phyloseq

## Preparation
```{r preparing, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
#Let's extract our taxa data frame
taxa <- short[,c(1,2,41:48)]
rownames(taxa) <- taxa$ASV
rownames(environ) <- environ$Samples

#Phyloseq DOES NOT LIKE dataframes (most of the times), USE MATRICES
taxa <- as.matrix(taxa)
lmabun <- t(abund) %>% as.data.frame()
lmabun <- add_column(lmabun, Total = rowSums(lmabun[c(1:34)])) %>% as.matrix()
```


## Phyloseq object
```{r phyloseq, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
OTU <- otu_table(lmabun, taxa_are_rows = TRUE)
TAXA <- tax_table(taxa)
SAMPLE <- sample_data(environ)
PHYLUM <- ape::as.phylo(clust)

remember <- phyloseq(OTU, TAXA, SAMPLE, PHYLUM)

```


## Phyloseq manipulation and explorarion
```{r explorer, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
plot_richness(remember, color = "Date", x = "Site")
plot_bar(remember, fill = "Division")
```


# Ploting relative abundance
One of the simplest ways to visualize data is to use ggplot, but in order to do so it is important to understand how and why we need to manipulate data formats. In this instance we are going to plot the relavite abundance of our data using a legacy labmicmar function and the transforming the object to get the wanted result.
```{r plot, echo=TRUE,warning=FALSE,message=FALSE,include=TRUE}
source("/Documents/Proyectos_R/Remember-R/Code/taxa_sum.R")
clase <- taxa.sum(abund,taxa,6) %>% as.data.frame()
rltv.class <- clase[,-19] %>% rltv.Otu.Table() %>% as.data.frame()
#rowSums(rltv.class)
rltv.class <- add_column(rltv.class, Sample = rownames(rltv.class), .before = "Mamiellophyceae", .name_repair = "minimal")
rltv.class$Sample <- factor(rltv.class$Sample, levels = rltv.class$Sample)
long <- reshape2::melt(rltv.class, id.vars = "Sample", variable.name = "Class")
grapho <- ggplot(long, aes(x=Sample,
                           y=value,
                           fill=Class)) + 
  geom_bar(position = "fill", stat = "identity") + 
  labs(x="Sampled sites", 
       y="Relative reads abundance",
       title = "Relative reads abundance at class level for analyzed ASV") + 
  scale_fill_viridis(discrete = TRUE, end = 0.89, direction = -1, option = "viridis") + 
  hrbrthemes::theme_ipsum(base_size = 22, axis_col = "#000000", axis_title_size = 30, plot_title_size = 30) + 
  theme(legend.position = "bottom", legend.direction = "horizontal")
legend <- cowplot::get_legend(grapho)
grapho <- grapho + theme(legend.position = "none")
ggsave(plot = grapho, 
       filename = "Figura_1.png",
       path = "/Documents/Proyectos_R/Remember-R/Products/", 
       width = 31, 
       height = 19, 
       dpi = 600)
ggsave(ggpubr::as_ggplot(legend), 
       filename = "Figura_1_legend.tiff", 
       path = "/Documents/Proyectos_R/Remember-R/Products/", 
       width = 21,
       height = 10,
       dpi = 900)
```

Agregar un plot de abundancia relativa