---
title: "Remember-R"
author: "Camilo Gálvez A."
output: html_document
---

# Libraries
```{r Libraries, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
library(tidyverse)
library(vegan)
library(clustsig)
library(Biostrings)
library(DECIPHER)
library(phyloseq)
library(ape)
library(dendextend)
library(heatmaply)
library(viridis)
library(htmlwidgets)
library(taxa)
library(metacoder)
library(ggvegan)
library(doParallel)
library(hrbrthemes)
```


# Data
```{r Data, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
data <- read.csv("https://raw.githubusercontent.com/cmlglvz/Remember-R/main/Data/abundancia_taxonomia.csv", header = TRUE, sep = ";", skip = 0, row.names = 1)
environ <- read.csv("https://raw.githubusercontent.com/cmlglvz/Remember-R/main/Data/environment.csv", header = TRUE, sep = ";", dec = ".", skip = 0, row.names = 1)
```


# Abundance: Raw, relative and transformed
```{r Abundance, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
short <- dplyr::filter(data,Total >= 20)
abund <- short[,c(3:36)] %>% t() %>% as.data.frame()
colnames(abund) <- short$ASV

#Benjamin Glasner
rltv.Otu.Table <- function(x) {
  x.Data.rltv <- NULL
  for (i in 1:dim(x)[1]) {
    x.Data.rltv <- rbind(x.Data.rltv, x[i,]/apply(x, 1, function(x) sum(x))[i])
  }
  rownames(x.Data.rltv) <- rownames(x)
  invisible(x.Data.rltv)
}

rltv <- rltv.Otu.Table(abund) %>% as.data.frame()
rowSums(rltv)

vrltv <- vegan::decostand(abund, method = "total")
rowSums(vrltv)

hell <- vegan::decostand(abund, method = "hellinger")
rowSums(hell)

#vegan::decostand(vrltv, method = "hellinger") %>% rowSums()
```


# SIMPROF
The similarity profile routine ([SIMPROF](https://sites.google.com/site/mb3gustame/hypothesis-tests/simprof); Clarke et al. 2008) is a hypothesis test which may be used during exploratory analyses. SIMPROF tests data, with no required a priori groupings (in contrast to e.g. ANOSIM), against the null hypothesis of "absence of structure". 
```{r SIMPROF, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
bray <- vegan::vegdist(hell, method = "bray", diag = TRUE) %>% as.matrix() %>% as.data.frame()
smprf <- clustsig::simprof(data = bray, method.cluster = "average", silent = FALSE, increment = 1000, undef.zero = TRUE)
simprof.plot(smprf)

#ver si esto es util
hellprof <- add_column(hell, 
                       .before = hell[,1], 
                       Site = c(rep("Site.1", 12),
                                rep("Site.3", 12), 
                                rep("Site.5", 10)
                                ), 
                       .name_repair = "minimal")
hellprof$Site <- factor(hellprof$Site, levels = unique(hellprof$Site))
```


# Sequences analysis
```{r FASTA, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
#nms <- paste(data$ASV, data$Species, sep = "_")
fasta <- data.frame(names = short$ASV, Sequences = all_of(short$Seq))
seqRFLP::dataframe2fas(fasta, "X:/Documents/Proyectos_R/Remember-R/Products/asv.fasta")
#Less than 1000 seqs = muscle.exe -align asv.fasta -output alignment.fasta
#More than 1000 seqs = muscle.exe -super5 asv.fasta -output alignment.afa
#it took 20 minutes in this case
```


# Alignment dendrogram
```{r alignment, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
multialign <- "/Documents/Proyectos_R/Remember-R/Products/alignment.fasta"
dbConn <- dbConnect(SQLite(), ":memory:")
Seqs2DB(multialign, type = "FASTA", dbFile = dbConn, "")
x <- dbGetQuery(dbConn, "select description from Seqs")$description
Add2DB(myData = data.frame(identifier = x, stringsAsFactors = FALSE), dbConn)
consensus <- IdConsensus(dbConn, threshold = 0.3, minInformation = 0.1)
distance.matrix <- DistanceMatrix(consensus, correction = "Jukes-Cantor", processors = NULL, verbose = TRUE)
dendrogram <- IdClusters(distance.matrix, method = "ML", showPlot = TRUE, type = "dendrogram", myXStringSet = consensus, processors = NULL, verbose = TRUE)
#Selected model was:TN93 + G4, it took 144.29 secs
dbDisconnect(dbConn)
```


# Cluster manipulation
```{r cluster, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
clust <- as.dendrogram(as.hclust(dendrogram)) %>% dendextend::set("branches_lwd", 0.3) %>% dendextend::ladderize(right = TRUE)
plot(clust)
sclust <- smprf$hclust
sclustS <- as.dendrogram(sclust) %>% set("branches_lwd", 0.3) %>% ladderize(right = FALSE)
plot(sclustS)
```


# Heatmap
```{r heatmap, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
maligned <- readDNAMultipleAlignment(multialign, format = "fasta")
mstring <- as(maligned, "DNAStringSet")
DECIPHER::BrowseSeqs(mstring, htmlFile = "/Documents/Proyectos_R/Remember-R/Products/multiple_alignment.html")

#Check if ID from .fasta and abundance table match
print(all(colnames(hell)%in%gtools::mixedsort(as.vector(rownames(as.matrix(maligned))), decreasing = FALSE)))
#If TRUE = they match

heatmap <- heatmaply(normalize(hell), 
                     Colv = clust, 
                     Rowv = sclustS, 
                     main = "Here goes the title", 
                     color = viridis(n = 256, 
                                     alpha = 1, 
                                     begin = 0, 
                                     end = 1, 
                                     option = "rocket")
                     )
htmlwidgets::saveWidget(heatmap, file = "/Documents/Proyectos_R/Remember-R/Products/heatmap.html")

wos.heat <- heatmaply(normalize(hell), 
                      Colv = clust, 
                      Rowv = NA, 
                      main = "Heatmap but without SIMPROF", 
                      color = viridis(n = 256, 
                                      alpha = 0.7, 
                                      begin = 0, 
                                      end = 1, 
                                      option = "mako")
                      )

saveWidget(wos.heat, file = "/Documents/Proyectos_R/Remember-R/Products/wos_heat.html")
```


# Using phyloseq
This next "chapter" is based on Daniel Vaulot's online tutorial for phyloseq

## Preparation
```{r preparing, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
#Let's extract our taxa data frame
taxa <- short[,c(1,2,41:48)]
rownames(taxa) <- taxa$ASV
rownames(environ) <- environ$Samples

#Phyloseq DOES NOT LIKE dataframes (most of the times), USE MATRICES
taxa <- as.matrix(taxa)
lmabun <- t(abund) %>% as.data.frame()
lmabun <- add_column(lmabun, Total = rowSums(lmabun[c(1:34)])) %>% as.matrix()
```


## Phyloseq object
```{r phyloseq, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
OTU <- otu_table(lmabun, taxa_are_rows = TRUE)
TAXA <- tax_table(taxa)
SAMPLE <- sample_data(environ)
PHYLUM <- ape::as.phylo(clust)

remember <- phyloseq(OTU, TAXA, SAMPLE, PHYLUM)

```


## Phyloseq manipulation and explorarion
```{r explorer, echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
plot_richness(remember, color = "Date", x = "Site")
plot_bar(remember, fill = "Division")
```

Agregar un plot de abundancia relativa